---
title: "pipeline_filtering_counts"
output: html_document
date: "2024-09-25"
params:
  code_dir: "None" # #to make compatible with R file
---

# To prepare a clean and final HIMSS dataset for analysis, our pipepline applies a series of logical filters. These filters remove entries for contacts with uncertain data, and leave only those which we are confident have quality entries. At the end of the pipeline, the final dataset therefore has far fewer entries.This script counts the number of observations in all the interim files produced throughout the pipeline filtering. 

```{r manual set up}
#User entry required!

#clear environment
rm(list = setdiff(ls(), c("params"))) # #to make compatible with R file

# *** IMPORTING FILES ***
# Add file names for all dataset outputs throughout the build pipeline

file_entities_contacts_0517 <- c("himss_entities_contacts_0517_v1.feather")

file_r_confirmed <- c("r_confirmed.feather")
file_r_remaining <- c("r_remaining.feather")

file_py_confirmed <- c("py_confirmed.csv")
file_py_remaining <- c("py_remaining.csv")

file_final_himss <- c("final_himss.feather")
file_final_confirmed <- c("final_confirmed.dta")

```

```{r setup, include=FALSE}
# NO manual entry required. This automatically loads necessary libraries and detects file paths.

# Install and load the here package, this will help auto detect the code directory and find the config.R script
if (!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
library(here)

# Optionally set the root directory manually (if needed)
# here::i_am("build_code/config.R")

# Construct the path to config.R
config_path <- here("build_code", "config.R")

# Print the config_path for debugging
print(paste("Config path:", config_path))

# Check if the config.R file exists
if (!file.exists(config_path)) {
  stop("The config.R file does not exist at the path: ", config_path)
}

# Source the config file
source(config_path)

# Now all necessary libraries are loaded, and file paths are set

# Clean up
rm(config_path)

```

```{r load data}

entities_contacts_0517 <- read_feather(paste0(derived_data,"/",file_entities_contacts_0517))

r_confirmed <- read_feather(paste0(auxiliary_data,"/",file_r_confirmed))
r_remaining <- read_feather(paste0(auxiliary_data,"/",file_r_remaining))

py_confirmed <- read_csv(paste0(auxiliary_data,"/",file_py_confirmed))
py_remaining <- read_csv(paste0(auxiliary_data,"/",file_py_remaining))

final_himss <- read_feather(paste0(derived_data,"/",file_final_himss))
final_confirmed <- read_dta(paste0(derived_data,"/",file_final_confirmed))
```


```{r function to count obs for each df}

count_hospital_rds <- function(df, col) {
  # Ensure the data frame is ungrouped
  df <- df %>% ungroup()
  
  # Create a grouping variable 'hospital_group' based on the presence of "Hospital" or "RDS" in the specified column
  df <- df %>%
    mutate(
      hospital_group = if_else(
        str_detect({{ col }}, "Hospital|RDS"),
        "Hospital/RDS",
        "Non-Hospital/RDS",
        missing = "Non-Hospital/RDS"  # Handle NA values by assigning them to "Non-Hospital/RDS"
      )
    )
  
  # Group by 'hospital_group' and summarize counts
  summary_counts <- df %>%
    group_by(hospital_group) %>%
    summarise(
      total_obs = n(),
      unique_contact_uniqueids = n_distinct(contact_uniqueid)
    ) %>%
    ungroup()
  
  # Get total count of observations in the entire data frame
  total_count <- nrow(df)
  
  # Return the summary_counts data frame and total_count
  list(
    summary_counts = summary_counts,
    total_count = total_count
  )
}

```

```{r apply function to all dataframes}
# List of all data frames to process
dfs <- list(
  entities_contacts_0517 = entities_contacts_0517,
  r_confirmed = r_confirmed,
  # r_remaining = r_remaining,
  py_confirmed = py_confirmed,
  # py_remaining = py_remaining,
  final_himss = final_himss
)

# Apply the updated count_hospital_rds function to all data frames
results <- map(dfs, ~count_hospital_rds(.x, entity_type))

# Extract and combine the summary counts from each data frame
summary_counts_list <- map2(results, names(results), ~mutate(.x$summary_counts, dataframe = .y))

# Combine all summary counts into a single data frame
combined_summary_counts <- bind_rows(summary_counts_list)

# View the combined summary counts
print(combined_summary_counts)

```


```{r plot the counts}
# Reshape the data to long format
combined_summary_counts_long <- combined_summary_counts %>%
  pivot_longer(
    cols = c(total_obs, unique_contact_uniqueids),
    names_to = "count_type",
    values_to = "count"
  )

# Recode count_type for better labels
combined_summary_counts_long <- combined_summary_counts_long %>%
  mutate(
    count_type = recode(
      count_type,
      total_obs = "Total Observations",
      unique_contact_uniqueids = "Unique # of Contacts"
    )
  )

# Remove NA values in 'count'
combined_summary_counts_long <- combined_summary_counts_long %>%
  filter(!is.na(count))

# Define a custom function to format numbers
format_count <- function(x) {
  ifelse(
    x >= 1e6, paste0(round(x / 1e6, 1), "M"),
    ifelse(
      x >= 1e3, paste0(round(x / 1e3, 1), "K"),
      as.character(x)
    )
  )
}

# Updated plot code applying custom formatting only to axis labels
ggplot(combined_summary_counts_long, aes(x = reorder(dataframe, count), y = count)) +
  geom_bar(stat = "identity", fill = "#1f77b4") +
  coord_flip(clip = "off") +
  geom_text(aes(label = comma(count)), hjust = -0.1, size = 2.5) +
  scale_y_continuous(labels = format_count, expand = expansion(mult = c(0, 0.2))) +
  labs(
    title = "Build Pipeline Filtering Funnel",
    x = "Dataframe",
    y = "Count"
  ) +
  facet_grid(
    count_type ~ hospital_group,
    scales = "free_y"
  ) +
  theme_minimal() +
  theme(
    plot.margin = unit(c(1, 3, 1, 1), "cm"),  # Adjust margins as needed
    panel.spacing = unit(2, "lines"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line.x = element_line(color = "black"),
    axis.line.y = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    # Adjust facet label text
    strip.text.y = element_text(
      size = 10,
      angle = 0,        # Keep labels horizontal
      hjust = 1,        # Right-align the text
      face = "bold",
      margin = margin(l = 10)
    ),
    strip.text.x = element_text(size = 10),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 8)
  )


```

